All Your Graphs Belong to Us

### 2012-07-17 doc/asplos13/implementation.tex
initial communication implementation section

Each scheduler has three queues:
\begin{description}
\item{Ready worker queue} This is a FIFO queue of tasks that are
  matched with workers and are ready to execute.
\item{Private task queue} This is a FIFO queue of tasks that must run on this core.
\item{Public task queue} This is a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool; this is described in the next section.
\end{description}

@@

\TODO{Jacob}

\subsection{Aggregation}

\subsection{Synchronization}

\subsection{Consistency}

>>>>

Each scheduler has three queues:
\begin{description}
\item{\bf Ready worker queue} This is a FIFO queue of tasks that are
  matched with workers and are ready to execute.
\item{\bf Private task queue} This is a FIFO queue of tasks that must run on this core.
\item{\bf Public task queue} This is a LIFO queue of tasks that are
  waiting to be matched with workers. It is a local partition of a shared
  task pool; this is described in the next section.
\end{description}

@@

In order to mitigate the low message injection rate limits of commodity
networks, Grappa's communication has two layers.

At the upper layer, Grappa implements asynchronous active messages
\cite{vonEicken92}. Each message consists of a function pointer, an
optional argument payload, and an optional data payload. When a task
sends a message, the message is copied to a send queue associated with
the message's destination and the task continues execution.

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. In our applications we see that the bulk of
the messages sent are small. \TODO{include stats} Commodity networks,
even ones designed for high-performance computing, provide low
bandwidth utilization at small message sizes. \TODO{include stats?} To
make the best use of the network, we must convert our small messages
into large ones.

Grappa leverages its lightweight context switching to do this
conversion. When a task sends a message, the message is not
immediately sent. Instead, it is stored in a queue associated with its
destination, to be sent later. If the task requires a response, it
suspends itself and the scheduler finds other work to run while the
task is waiting. When the reply arrives, the suspended task is woken.

There are three situations in which a queue of aggregated messages is
sent. First, each queue has a message size threshold, chosen to give
reasonable network performance. If the size in bytes of a queue is
above the threshold, the contents of the queue are sent
immediately. Second, each queue has a wait time threshhold. If the
oldest message in a queue has been waiting longer than this threshold,
the contents of the queue are sent immediately, even if the queue size
is lower than the message size threshold.  Third, queues may be
explicitly flushed in situations where the programmer wants to
minimize the latency of a message at the cost of bandwidth
utilization.

The network layer is serviced by polling. Periodically when a context
switch occurs, the Grappa scheduler switches to the network polling
thread. This thread has three responsibilities. First, it polls the
lower-level network layer to ensure it makes progress. Second, it
deaggregates received messages and executes active message
handlers. Third, it checks to see if any aggregation queues have
messages that have been waiting longer than the threshold; if so, it
sends them.

Underneath the aggregation layer, Grappa uses the GASNet communication
library \cite{GASNet} to actually move data. All interprocess
communication, whether on or off a cluster node, is handled by the
GASNet library. GASNet is able to take advantage of many communication
mechanism, including Ethernet and Infiniband between nodes, as well as
shared memory within a node.

Some networks provide access to a remote machine's memory
directly. This would seem to be a good fit for a programming model
focused on global shared memory, but in fact we do not use it. RDMA
operations are subject to the same message rate limitations as all
other messages on these cards, and thus using raw RDMA operations for
our small messages would make inefficient use of bandwidth. Instead,
we implement remote memory operations with active messages as
described in the next section. This means that Grappa is not limited
to RDMA-capable networks.

\subsection{Memory}

remote access

heap

allocation

synchronization

consistency
>>>>>>> Stashed changes
